---
title: "Bankruptcy Analysis"
author: "Mandy HP Nguyen"
date: "12/17/2021"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    theme: paper
    highlight: breezedark
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: yes
  word_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Environment Setting

## Load Mandy's Functions

```{r}
source("https://raw.githubusercontent.com/mandyhpnguyen/Mandy-Functions/main/M.R-Funcs/general.R")
```

## Load Packages

```{r}
pkgs <- c(
  "beepr",
  # General pkgs:
  "tidyverse", "dplyr", "summarytools", "reshape2", "pastecs", "ROSE",
  # Analytics pkgs:
  "forecast", "zoo", "scorecard",
  # Evaluation pkgs:
  "caret", "DT", "lift", "gains", "gt", "cvms","tibble", "fourfoldplot",
  # Visualization pkgs:
  "RColorBrewer", "hrbrthemes", "knitr", "showtext", "sysfonts",
  "ggfortify", "ggplot2", "corrplot", "GGally", "viridis",
  "lattice", "grid", "cowplot", "Amelia"
)
suppressMessages(suppressWarnings(loadpkg(pkgs)))
```

## Set fonts

```{r}
# Load the main font I used in my paper to plot charts for standardization
windowsFonts(cambria = windowsFont("Cambria"))
my.par <- function() {
  options(scipen = 999)
  par(mfrow = c(1, 1),
      family = "cambria", 
      cex.main = 1.25, cex.lab = 1.25, cex.axis = 1.25)
}
```

## Color Scheme

```{r}
pal <- c("cornflowerblue", "orange", "red2", "forestgreen", "mediumorchid1", "tomato3", "cadetblue1")
```

# Collect Data

```{r}
book <- read.csv("https://raw.githubusercontent.com/mandyhpnguyen/MS-Data-Analytics-Datasets/main/Data%20Mining/CharlesBookClub.csv")
```

# Pre-Process Data

```{r}
cbc <- book
```

## Compute RFM_score

```{r}
cbc$RFM_score <- paste(cbc$Rcode, cbc$Fcode, cbc$Mcode)
cbc$RFM_score <- gsub(" ", "", cbc$RFM_score)
cbc$RFM_score <- as.factor(cbc$RFM_score)
cbc[1:10, "RFM_score"]
```

## Factor Categorical Attributes

```{r}
cbc[, "Gender"] <- factor(cbc[, "Gender"])
cbc[, "Yes_Florence"] <- factor(cbc[, "Yes_Florence"], levels = c(1, 0))
cbc[, "No_Florence"] <- factor(cbc[, "No_Florence"], levels = c(0, 1))

cbc[, "Mcode"] <- factor(cbc[, "Mcode"])
cbc[, "Rcode"] <- factor(cbc[, "Rcode"])
cbc[, "Fcode"] <- factor(cbc[, "Fcode"])
```

# Parition

```{r}
set.seed(2021)
train.index <- sample(1:nrow(cbc), 0.7*nrow(cbc))
train <- cbc[train.index, ]
valid <- cbc[-train.index, ]
```

# RFM Analysis

## Compute Response Rate

```{r}
# Response rate for training data
# Fix height issue: https://stackoverflow.com/questions/16121903/r-barplot-y-axis-scale-too-short
rr_bar <- function(X, Z=""){
  barplot(height = X, space = .2,
          horiz = FALSE, density = 50, angle = 60,
          col = c("red2", "cornflowerblue"), border = NA,
          ylim = range(pretty(c(0, X))),
          xlab = 'Florence',
          ylab = 'Frequency of Response Rate',
          main = Z
  ) %>% 
    text(0, y = X, labels = paste0(round(X*100,2), "%"),
         cex = 1, pos = 3)
}
```

```{r}
my.par() %>% par(mfrow = c(1, 3))
rr_bar(prop.table(table(cbc$Florence)), 
                   "Raw Set")
rr_bar(prop.table(table(train$Florence)), 
                   "Training Set")
rr_bar(prop.table(table(valid$Florence)),
                   "Validation Set")
```

### Response rate of Whole Data Set

```{r}
toprr_cbc <- cbc %>% 
  group_by(RFM_score) %>% 
  summarise(Yes = length(Yes_Florence[Yes_Florence == '1']),
            No = length(No_Florence[No_Florence == '1']),
            Response_Rate = round(Yes / (Yes + No), 3),
            Count = length(RFM_score)
  ) %>% 
  arrange(desc(Response_Rate)) %>% 
  as.data.frame()
toprr_cbc
```
```{r}
setwd("C:/One Drives/OneDrive - Webster University/Webster Classes/21F_CSDA 6010_Practicum")
options(scipen = 0, digits = 2)
toprr_cbc %>% write.csv("toprr_cbc.csv", row.names = TRUE)
```

### Response Rate of Training

```{r}
toprr_train <- train %>% 
  group_by(RFM_score) %>% 
  summarise(Yes = length(Yes_Florence[Yes_Florence == '1']),
            No = length(No_Florence[No_Florence == '1']),
            Response_Rate = round(Yes / (Yes + No), 3),
            Count = length(RFM_score)
  ) %>% 
  arrange(desc(Response_Rate)) %>% 
  as.data.frame()
```

```{r}
setwd("C:/One Drives/OneDrive - Webster University/Webster Classes/21F_CSDA 6010_Practicum")
options(scipen = 0, digits = 2)
toprr_train %>% write.csv("toprr_train.csv", row.names = TRUE)
```

### Response Rate of Validation

```{r}
toprr_valid <- valid %>% 
  group_by(RFM_score) %>% 
  summarise(Yes = length(Yes_Florence[Yes_Florence == '1']),
            No = length(No_Florence[No_Florence == '1']),
            Response_Rate = round(Yes / (Yes + No), 3),
            Count = length(RFM_score)
  ) %>% 
  arrange(desc(Response_Rate)) %>% 
  as.data.frame()
```

```{r}
setwd("C:/One Drives/OneDrive - Webster University/Webster Classes/21F_CSDA 6010_Practicum")
options(scipen = 0, digits = 2)
toprr_valid %>% write.csv("toprr_valid.csv", row.names = TRUE)
```

## Evaluate with different cut-offs

## Generic Functions

```{r}
RFM_emore <- function(cut_off, equation="") {
  cat(">> Valid >= Cut-off =", cut_off, "\n",
      ifelse(equation == "", "", paste0("> Equation: ", equation)), "\n\n")
  train_c <- toprr_train
  train_c$Response <- factor(ifelse(train_c$Response_Rate >= cut_off, "Yes", "No"))
  valid_c <- toprr_valid
  valid_c$Response_V <- factor(ifelse(valid_c$Response_Rate >= cut_off, "Yes", "No"))
  index1 <- valid_c$RFM_score
  valid_c$Response_T <- train_c[index1, ]$Response
  valid_c <- valid_c[, c(1, 4:7)]
  confusionMatrix(valid_c$Response_T, valid_c$Response_V)
}
```

```{r}
RFM_eless <- function(cut_off, equation="") {
  cat(">> Valid <= Cut-off =", cut_off, "\n",
      ifelse(equation == "", "", paste0("> Equation: ", equation)), "\n\n")
  train_c <- toprr_train
  train_c$Response <- factor(ifelse(train_c$Response_Rate <= cut_off, "Yes", "No"))
  valid_c <- toprr_valid
  valid_c$Response_V <- factor(ifelse(valid_c$Response_Rate <= cut_off, "Yes", "No"))
  index1 <- valid_c$RFM_score
  valid_c$Response_T <- train_c[index1, ]$Response
  valid_c <- valid_c[, c(1, 4:7)]
  confusionMatrix(valid_c$Response_T, valid_c$Response_V)
}
```

## Means

```{r}
mean(toprr_cbc$Response_Rate)
mean(toprr_train$Response_Rate)
```

## More Comparison

```{r}
more1 <- RFM_emore(mean(toprr_train$Response_Rate), "mean(toprr_train$Response_Rate)")
more2 <- RFM_emore(mean(toprr_train$Response_Rate)*2, "mean(toprr_train$Response_Rate)*2")
more3 <- RFM_emore(mean(toprr_cbc$Response_Rate), "mean(toprr_cbc$Response_Rate)")
more4 <- RFM_emore(mean(toprr_cbc$Response_Rate)*2, "mean(toprr_cbc$Response_Rate)*2")
```

## Less Comparison

```{r}
less1 <- RFM_eless(mean(toprr_train$Response_Rate), "mean(toprr_train$Response_Rate)")
less2 <- RFM_eless(mean(toprr_train$Response_Rate)*2, "mean(toprr_train$Response_Rate)*2")
less3 <- RFM_eless(mean(toprr_cbc$Response_Rate), "mean(toprr_cbc$Response_Rate)")
less4 <- RFM_eless(mean(toprr_cbc$Response_Rate)*2, "mean(toprr_cbc$Response_Rate)*2")
```

## CM Results

```{r}
more1
more2
more3
more4
less1
less2
less3
less4
```


## CM Plots

```{r}
rfm_cmp <- function(cm, title) {
  fourfoldplot(cm, color = c("cornflowerblue", "red2"),
               margin = 1, space = 0.3,
               conf.level = 0, main = title)
}
my.par() %>% par(mfrow = c(2, 4),
                 mai = par("mai") * 1,
                 oma = c(0,0,0,0) + 0.1,
                 mar = c(2,4,2,0) + 0.1
                 )
rfm_cmp(more1$table, "A-1")
rfm_cmp(more2$table, "A-2")
rfm_cmp(more3$table, "A-3")
rfm_cmp(more4$table, "A-4")
rfm_cmp(less1$table, "A-5")
rfm_cmp(less2$table, "A-6")
rfm_cmp(less3$table, "A-7")
rfm_cmp(less4$table, "A-8")
```


# Clean up

```{r}
rm(list = ls())
cat("\014")
```
